
```{r}
library(Rcpp)
library(dada2)
library(phyloseq)
library(permute)
library(lattice)
library(vegan)
library(ggplot2)
library(tidyverse)
library(ggstatsplot)
library(dplyr)
library(microbiome)
library(microbiomeutilities) 
library(knitr)
library(RColorBrewer)
library(DT)
library(gt)
library(cowplot)
library(PerformanceAnalytics)


```


```{r}
ps.prev <- readRDS("phyobjects/STEP 15/ps.prev.f.rds")
ps_tss_1 <- transform_sample_counts(ps.prev, function(x){x / sum(x)})
mtd <- data.frame(sample_data(ps.prev), check.names = FALSE)
```


```{r}
# compute OTU prevalence
prvl <- list()
for (i in c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV", 
            "water.MMV", "water.VMV")) {
    prvl[[i]] <- subset_samples(ps_tss_1, region_regime == i) %>%
      prevalence(detection = 0.0005, count = TRUE, include.lowest = TRUE)
}
for (i in c("feed.M", "feed.V")) {
    prvl[[i]] <- subset_samples(ps_tss_1, region_regime == i) %>%
      prevalence(detection = 0.0005, count = TRUE, include.lowest = TRUE)
}
```


Compute ASV overlap

```{r}
ovrl <- bind_cols(prvl) %>%
  pivot_longer(pyloric.MMV:distal.VMV, 
               names_to = "SAMPLE", 
               values_to = "GUT") %>%
  # compute OTU overlap between water/feed and intestinal samples
  mutate(
    WaterOverlappyloric.MMV = case_when(
      water.MMV > 0 & GUT >  0 ~ "Shared",
      water.MMV > 0 & GUT == 0 ~ "Unique to water",
      water.MMV == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    WaterOverlappyloric.VMV = case_when(
      water.VMV > 0 & GUT >  0 ~ "Shared",
      water.VMV > 0 & GUT == 0 ~ "Unique to water",
      water.VMV == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    WaterOverlapmiddle.MMV = case_when(
      water.MMV > 0 & GUT >  0 ~ "Shared",
      water.MMV > 0 & GUT == 0 ~ "Unique to water",
      water.MMV == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    WaterOverlapmiddle.VMV = case_when(
      water.VMV > 0 & GUT >  0 ~ "Shared",
      water.VMV > 0 & GUT == 0 ~ "Unique to water",
      water.VMV == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    WaterOverlapdistal.MMV = case_when(
      water.MMV > 0 & GUT >  0 ~ "Shared",
      water.MMV > 0 & GUT == 0 ~ "Unique to water",
      water.MMV == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    WaterOverlapdistal.VMV = case_when(
      water.VMV > 0 & GUT >  0 ~ "Shared",
      water.VMV > 0 & GUT == 0 ~ "Unique to water",
      water.VMV == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    FeedOverlappyloric.MMV = case_when(
      feed.V > 0 & GUT >  0 ~ "Shared",
      feed.V > 0 & GUT == 0 ~ "Unique to feed",
      feed.V == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    FeedOverlappyloric.VMV = case_when(
      feed.V > 0 & GUT >  0 ~ "Shared",
      feed.V > 0 & GUT == 0 ~ "Unique to feed",
      feed.V == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    FeedOverlapmiddle.MMV = case_when(
      feed.V > 0 & GUT >  0 ~ "Shared",
      feed.V > 0 & GUT == 0 ~ "Unique to feed",
      feed.V == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    FeedOverlapmiddle.VMV = case_when(
      feed.V > 0 & GUT >  0 ~ "Shared",
      feed.V > 0 & GUT == 0 ~ "Unique to feed",
      feed.V == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    FeedOverlapdistal.MMV = case_when(
      feed.V > 0 & GUT >  0 ~ "Shared",
      feed.V > 0 & GUT == 0 ~ "Unique to feed",
      feed.V == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent"),
    FeedOverlapdistal.VMV = case_when(
      feed.V > 0 & GUT >  0 ~ "Shared",
      feed.V > 0 & GUT == 0 ~ "Unique to feed",
      feed.V == 0 & GUT >  0 ~ "Unique to gut",
      TRUE ~ "Absent")) %>%
  pivot_longer(WaterOverlappyloric.MMV:FeedOverlapdistal.VMV, 
               names_to = "OverlapType", 
               values_to = "Category") %>%
  count(OverlapType, SAMPLE, Category) %>%
  mutate(
    Category = factor(
      Category, 
      levels = c("Unique to water", "Unique to feed", 
                 "Shared", "Unique to gut", "Absent")),
    #Diet = ifelse(grepl("FM", Sample), "FM", "ICJ", "ACJ", "IWA", "AWA", "SBM"),
    Region_Regime = factor(SAMPLE, levels = c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV")))

```

Plotting
Plot microbial overlap between feed and intestinal samples.

```{r}
ovrl_bar_feed <- ovrl %>%
  filter(OverlapType == "FeedOverlappyloric.MMV" & Region_Regime == "pyloric.MMV" |
         OverlapType == "FeedOverlappyloric.VMV"  & Region_Regime == "pyloric.VMV" | 
         OverlapType == "FeedOverlapmiddle.MMV"  & Region_Regime == "middle.MMV"| 
         OverlapType == "FeedOverlapmiddle.VMV"  & Region_Regime == "middle.VMV" | 
         OverlapType == "FeedOverlapdistal.MMV"  & Region_Regime == "distal.MMV"| 
         OverlapType == "FeedOverlapdistal.VMV"  & Region_Regime == "distal.VMV") %>%
  filter(Category != "Absent") %>%
  ggplot(aes(x = SAMPLE, c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV"), 
             y = n, fill = Category, label = n)) +
  geom_bar(stat = "identity") +
  geom_text(position = position_stack(vjust = 0.5), color = "white", size = 2.5) +
  labs(x = "", y = "Number of ASVs", size = 10) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_manual(values = c("#B2DF8A", "grey50", "#33A02C")) +
  #facet_nested(~ Diet + Site, scales = "free_x", nest_line = TRUE) +
  scale_x_discrete(limit=c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV")) +
  theme_bw() +
  theme(axis.text.x = element_text (size = 9),
        strip.background = element_blank(),
        legend.title = element_blank(),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, -5)) +
  theme(axis.text.x.bottom = element_text(angle = 90)) 

```


Plot microbial overlap between water and intestinal samples.

```{r}
ovrl_bar_water <- ovrl %>%
  filter(OverlapType == "WaterOverlappyloric.MMV" & Region_Regime == "pyloric.MMV" |
         OverlapType == "WaterOverlappyloric.VMV"  & Region_Regime == "pyloric.VMV" | 
         OverlapType == "WaterOverlapmiddle.MMV"  & Region_Regime == "middle.MMV"| 
         OverlapType == "WaterOverlapmiddle.VMV"  & Region_Regime == "middle.VMV" | 
         OverlapType == "WaterOverlapdistal.MMV"  & Region_Regime == "distal.MMV"| 
         OverlapType == "WaterOverlapdistal.VMV"  & Region_Regime == "distal.VMV")  %>%
  filter(Category != "Absent") %>%
  ggplot(aes(x = SAMPLE, c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV"), 
             y = n, fill = Category, label = n)) +
  geom_bar(stat = "identity") +
  geom_text(position = position_stack(vjust = 0.5), color = "white", size = 2.5) +
  labs(x = "", y = "Number of ASVs", size = 10) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + 
  scale_fill_manual(values = c("#A6CEE3", "grey50", "#1F78B4")) +
  #facet_nested(~ Diet + Site, scales = "free_x", nest_line = TRUE) +
  scale_x_discrete(limit=c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV")) +
  theme_bw() +
  theme(axis.text.x = element_text (size = 9),
        strip.background = element_blank(),
        legend.title = element_blank(),
        legend.margin = margin(0, 0, 0, 0),
        legend.box.margin = margin(0, 0, 0, -5)) +
  theme(axis.text.x.bottom = element_text(angle = 90)) 
```

Assemble plots

```{r}
ovrl_bar_feed + ovrl_bar_water  
```

Abundance of ASVs shared

Compute shared ASV abundance

Compute the abundance of shared ASVs between feed and intestinal samples.


```{r}
shared_feed <- map(c("feed.V", "feed.V", "feed.V", "feed.V", "feed.V", "feed.V"), function(x) {
  prune_taxa(names(which(prvl[[x]] > 0)), ps_tss_1) %>% sample_sums()
  }
)
names(shared_feed) <- c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV")
# names(prvl)

shared_feed <- map(seq_along(shared_feed), function(x){
  data.frame(Abundance = shared_feed[[x]]) %>%
  rownames_to_column("sample_name") %>%
  inner_join(rownames_to_column(mtd, "sample_name"), by = "sample_name") %>%
  filter(Region_Regime == names(shared_feed)[x] & !sample %in% c("feed", "water")) %>%
  #mutate(Segment = ifelse(Segment == "PI", "PIM", "DIM"),
         #Segment = factor(Segment, levels = c("PIM", "DIM"))) %>%
  mutate(Abundance = round(100 * Abundance, 1))}) %>%
  bind_rows()
```


Compute the abundance of shared OTUs between water and intestinal samples.

```{r}
shared_water <- map(c("water.MMV", "water.VMV", "water.MMV", "water.VMV", "water.MMV", "water.VMV"), function(x) {
  prune_taxa(names(which(prvl[[x]] > 0)), ps_tss_1) %>% sample_sums()
}
)
names(shared_water) <- c("pyloric.MMV", "pyloric.VMV", "middle.MMV", "middle.VMV", "distal.MMV", "distal.VMV")
# names(prvl)

shared_water <- map(seq_along(shared_water), function(x){
  data.frame(Abundance = shared_water[[x]]) %>%
    rownames_to_column("sample_name") %>%
    inner_join(rownames_to_column(mtd, "sample_name"), by = "sample_name") %>%
    filter(Region_Regime == names(shared_water)[x] & !sample %in% c("feed", "water")) %>%
    #mutate(Segment = ifelse(Segment == "PI", "PIM", "DIM"),
    #Segment = factor(Segment, levels = c("PIM", "DIM"))) %>%
    mutate(Abundance = round(100 * Abundance, 1))}) %>%
  bind_rows()
```

Plotting
Abundance of shared OTUs between feed and intestinal samples.

```{r}
bp_shared_feed <- ggplot(shared_feed, aes(x = Region_Regime, y = Abundance)) +
  geom_boxplot(fill = "grey50", width = 0.5) +
  geom_point() +
  #facet_nested(~ Diet + Segment, scales = "free_x", nest_line = TRUE) +
  labs(x = "", y = "Relative abundance of shared ASVs (%)", size = 10) +
  #scale_x_discrete(limit=c("M", "V", "MM", "VM", "MMV", "VMV")) +
  scale_y_continuous(
    limits = c(0, 100), breaks = 0:10*10, expand = expansion(c(0, 0))
    ) + 
  theme_bw() +
  theme(axis.text.x = element_text (size = 9),
        strip.background = element_blank(),
        legend.position = "none") +
  theme(axis.text.x.bottom = element_text(angle = 90)) 

```

Abundance of shared OTUs between water and intestinal samples.

```{r}
bp_shared_water <- shared_water %>%
  ggplot(aes(x = Region_Regime, y = Abundance)) +
    geom_boxplot(fill = "grey50", width = 0.5) +
    geom_point() +
    #facet_nested(~ Diet + Segment, scales = "free_x", nest_line = TRUE) +
    labs(x = "", y = "Relative abundance of shared ASVs (%)", size = 10) +
    scale_y_continuous(
      limits = c(0, 100), breaks = 0:10*10, expand = expansion(c(0, 0))
      ) + 
    #scale_x_discrete(limit=c("M", "V", "MM", "VM", "MMV", "VMV")) +
    theme_bw() +
    theme(axis.text.x = element_text (size = 9),
          strip.background = element_blank(),
          legend.position = "none") +
  theme(axis.text.x.bottom = element_text(angle = 90)) 

```

Assemble plots
```{r}
bp_shared_feed + bp_shared_water
```

Figure 7

```{r}
library(patchwork)
fig7a <- ovrl_bar_feed + bp_shared_feed + 
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = list(c("A", ""))) &
  theme(
    legend.position = "top",
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(10, 10, 10, 10))
fig7b <- ovrl_bar_water + bp_shared_water +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = list(c("B", ""))) &
  theme(
    legend.position = "top",
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(10, 10, 10, 10))
wrap_elements(full = fig7a) | wrap_elements(full = fig7b) 

# export plot
ggsave(("figures/sharedASVs.RA.tiff"), width = 12, height = 5,
       units = "in", dpi = 300, compression = "lzw")

```
